---
title: "Summarize_indices"
author: "Kelly"
date: "2025-08-25"
output: html_document
---

```{r}
#0# Workspace opruimen
rm(list=ls())
```


```{r}
#load packages
library(sf)
library(purrr)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(writexl)
library(plotly)
library(scales)
library(ggnewscale)
library(patchwork)
library(sf)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
# Function to read shapefiles, compute indicators, summarise per city
library(tidyverse)
library(viridis)
```

#STEP 1: LOAD data

FLANDERS
```{r}
#load flanders

# path to folder
shp_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/BBK_15/Buffers"

# list shapefiles
shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)

# function to read shapefile and add city column
read_with_city <- function(file) {
  city <- str_extract(basename(file), "^[A-Za-z]+")  # take leading word (e.g. Aalst)
  st_read(file, quiet = TRUE) %>%
    mutate(city = city)
}

# read all and bind.
all_gardens_15 <- map(shp_files, read_with_city) %>%
  bind_rows()
```

#probleem met tuinen die op de rand van twee rasters vallen? of waarvan 400 m erbuiten niet meer binnen raster valt?
```{r}
#filter
all_gardens_15 <- all_gardens_15 %>% filter(ex50 != 0)

sum(duplicated(all_gardens_15$CAPAKEY))
```

```{r}
#calculate index (huidig)
all_gardens_15$UGEI_50h <- all_gardens_15$BAF1_50-all_gardens_15$risk50
all_gardens_15$UGEI_100h <- all_gardens_15$BAF1_100-all_gardens_15$risk100
all_gardens_15$UGEI_200h <- all_gardens_15$BAF1_200-all_gardens_15$risk200
all_gardens_15$UGEI_400h <- all_gardens_15$BAF1_400-all_gardens_15$risk400

#calculate index (potentieel)
all_gardens_15$UGEI_50p <- all_gardens_15$BAF2_50-all_gardens_15$risk50
all_gardens_15$UGEI_100p <- all_gardens_15$BAF2_100-all_gardens_15$risk100
all_gardens_15$UGEI_200p <- all_gardens_15$BAF2_200-all_gardens_15$risk200
all_gardens_15$UGEI_400p <- all_gardens_15$BAF2_400-all_gardens_15$risk400

all_gardens_15 <- all_gardens_15[complete.cases(all_gardens_15$exp50), ] 
```

MANCHESTER
```{r}
#load Manchester
Manchester_gardens <- st_read('C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/Manchester/Manchester.CL_Landcover.metrics.shp')
```

```{r}
#filter
Manchester_gardens <- Manchester_gardens %>% filter(ex50 != 0)
Manchester_gardens$city <- 'Manchester'
```

```{r}
#calculate index
Manchester_gardens$UGEI_50h <- Manchester_gardens$BAF1_50-Manchester_gardens$risk50
Manchester_gardens$UGEI_100h <- Manchester_gardens$BAF1_100-Manchester_gardens$risk100
Manchester_gardens$UGEI_200h <- Manchester_gardens$BAF1_200-Manchester_gardens$risk200
Manchester_gardens$UGEI_400h <- Manchester_gardens$BAF1_400-Manchester_gardens$risk400

Manchester_gardens$UGEI_50p <- Manchester_gardens$BAF2_50-Manchester_gardens$risk50
Manchester_gardens$UGEI_100p <- Manchester_gardens$BAF2_100-Manchester_gardens$risk100
Manchester_gardens$UGEI_200p <- Manchester_gardens$BAF2_200-Manchester_gardens$risk200
Manchester_gardens$UGEI_400p <- Manchester_gardens$BAF2_400-Manchester_gardens$risk400
```

#STEP 2: SUPPLY and DEMAND figures (descriptive)

```{r}
# --- Drop geometry and combine datasets ---
combined_data <- bind_rows(
  st_drop_geometry(all_gardens_15) %>% mutate(dataset = "Flanders"),
  st_drop_geometry(Manchester_gardens) %>% mutate(dataset = "Manchester")
)

# Function to create plot with vertical legend underneath
prepare_metric_plot_vert_legend <- function(df, metric_col, metric_name, breaks, labels) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  # Calculate percentages
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  # Palette
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- colorRampPalette(c("#c6d9e4", "#407ca2"))(n_cats)
  
  # Plot
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 5, color = "black") +
    scale_fill_manual(values = my_palette, name = "Level",
                      guide = guide_legend(direction = "horizontal")) +
    labs(x = NULL, y = "% of gardens", title = metric_name) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          legend.position = "bottom")
  
  return(p)
}

# Create all four plots
# Font size settings
base_size <- 24  # adjust as needed

breaks_labels <- list(
  BAF1_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("0-0.025", "0.025-0.05", "0.05-0.075", "0.075-0.1", ">0.1")
  ),
  BAF2_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("0-0.025", "0.025-0.05", "0.05-0.075", "0.075-0.1", ">0.1")
  ),
  BAFtot_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("0-0.025", "0.025-0.05", "0.05-0.075", "0.075-0.1", ">0.1")
  ),
  BAFper_50 = list(
    breaks = c(-Inf, 20, 40, 60, 80, +Inf),
    labels = c("0-20", "20-40", "40-60", "60-80", "80-100")
  )
)

# --- Plot 1 (skip legend) ---
p1 <- prepare_metric_plot_vert_legend(combined_data, "BAF1_50", "A. Garden Contribution",
                                      breaks_labels[["BAF1_50"]]$breaks,
                                      breaks_labels[["BAF1_50"]]$labels) +
      theme(
        plot.title = element_text(size = base_size-2, face = "bold"),  # title above each plot
        axis.title = element_text(size = base_size),
        axis.text = element_text(size = base_size),
        axis.text.x = element_text(size = base_size),
        legend.text = element_text(size = base_size - 4),
        legend.title = element_text(size = base_size - 2)
      ) +
      guides(fill = "none")  # skip legend

# --- Plot 2 (keep horizontal legend) ---
p2 <- prepare_metric_plot_vert_legend(combined_data, "BAF2_50", "B. Garden Potential",
                                      breaks_labels[["BAF2_50"]]$breaks,
                                      breaks_labels[["BAF2_50"]]$labels) +
      scale_fill_manual(values = colorRampPalette(c("#c6d9e4", "#407ca2"))(5),
                        name = "Levels",  # <- changed legend title
                        guide = guide_legend(direction = "horizontal", nrow = 1)) +
      theme(legend.position = "bottom")+
      theme(
        plot.title = element_text(size = base_size-2, face = "bold"),  # title above each plot
        axis.title = element_text(size = base_size),
        axis.text = element_text(size = base_size),
        axis.text.x = element_text(size = base_size),
        axis.title.y = element_blank(),   # remove y-axis title
        axis.text.y  = element_blank(),   # remove y-axis text
        axis.ticks.y = element_blank(),   # remove y-axis ticks
        legend.text = element_text(size = base_size - 4),
        legend.title = element_text(size = base_size- 2)
      ) 

# --- Plot 3 (skip legend) ---
p3 <- prepare_metric_plot_vert_legend(combined_data, "BAFtot_50", "C. PGI",
                                      breaks_labels[["BAFtot_50"]]$breaks,
                                      breaks_labels[["BAFtot_50"]]$labels) +
      guides(fill = "none")  + # skip legend
      theme(
        plot.title = element_text(size = base_size-2, face = "bold"),  # title above each plot
        axis.title = element_text(size = base_size),
        axis.text = element_text(size = base_size),
        axis.text.x = element_text(size = base_size),
        axis.title.y = element_blank(),   # remove y-axis title
        axis.text.y  = element_blank(),   # remove y-axis text
        axis.ticks.y = element_blank(),   # remove y-axis ticks
        legend.text = element_text(size = base_size - 4),
        legend.title = element_text(size = base_size- 2)
      ) 

# --- Plot 4 (keep horizontal legend) ---
p4 <- prepare_metric_plot_vert_legend(combined_data, "BAFper_50", "D. CEI",
                                      breaks_labels[["BAFper_50"]]$breaks,
                                      breaks_labels[["BAFper_50"]]$labels) +
      scale_fill_manual(values = colorRampPalette(c("#c6d9e4", "#407ca2"))(5),
                        name = "Levels",  # <- changed legend title
                        guide = guide_legend(direction = "horizontal", nrow = 2)) +
      theme(legend.position = "bottom") +
      theme(
        plot.title = element_text(size = base_size-2, face = "bold"),  # title above each plot
        axis.title = element_text(size = base_size),
        axis.text = element_text(size = base_size),
        axis.text.x = element_text(size = base_size),
        axis.title.y = element_blank(),   # remove y-axis title
        axis.text.y  = element_blank(),   # remove y-axis text
        axis.ticks.y = element_blank(),   # remove y-axis ticks
        legend.text = element_text(size = base_size - 4),
        legend.title = element_text(size = base_size- 2)
      ) 

# --- Combine horizontally ---
combined_plot <- (p1 | p2 | p3 | p4)
combined_plot

# --- Save figure ---
ggsave("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Figures/BAF_metrics_plot_horizontal_legends.png",
       plot = combined_plot,
       width = 16, height = 8, dpi = 300)
```

```{r}
# Example mapping: replace these with your actual city names
niscode_names <- c(
  "11002" = "Antwerp",
  "12025" = "Mechelen",
  "13040" = "Turnhout",
  "24062" = "Leuven",
  "31005" = "Bruges",
  "34022" = "Kortrijk",
  "35013" = "Ostend",
  "36015" = "Roeselare",
  "41002" = "Aalst",
  "44021" = "Ghent",
  "46021" = "Sint-Niklaas",
  "71016" = "Genk",
  "71022" = "Hasselt"
)

# --- Step 1: Filter Flanders only ---
flanders_gardens <- st_drop_geometry(all_gardens_15) %>%
  filter(NISCODE %in% names(niscode_names)) %>%   # keep only 13 cities
  mutate(dataset = as.factor(NISCODE))  # use NISCODE as dataset

flanders_gardens <- flanders_gardens %>%
  mutate(dataset = niscode_names[as.character(NISCODE)])

# --- Step 2: Plotting function ---
make_blue_palette <- function(n) {
  colorRampPalette(c("#c6d9e4", "#407ca2"))(n)  # light to dark red
}

base_size <- 26
# --- Function to create plots with horizontal legend (same as before, adapted) ---
prepare_metric_plot_vert_legend <- function(df, metric_col, metric_name, breaks, labels, show_legend = TRUE, show_y = FALSE) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- make_blue_palette(n_cats)
  
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 5, color = "black") +
    labs(x = NULL, y = if(show_y) "% of gardens" else NULL, title = metric_name) +
    scale_fill_manual(values = my_palette,
                      name = "Levels",
                      guide = if(show_legend) guide_legend(direction = "horizontal", nrow = 1) else "none") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = base_size-2, face = "bold"),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size),
      axis.text.x = element_text(angle = 30, hjust = 1, size = base_size),
      axis.title.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.text.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.ticks.y = if(show_y) element_line() else element_blank(),
      legend.text = element_text(size = base_size - 4),
      legend.title = element_text(size = base_size -2 ),
      legend.position = if(show_legend) "bottom" else "none"
    )
  
  return(p)
}
# --- Step 3: Define breaks/labels ---
breaks_labels <- list(
  BAF1_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAF2_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAFtot_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAFper_50 = list(
    breaks = c(-Inf, 20, 40, 60, 80, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  )
)

# Supply plots
p1 <- prepare_metric_plot_vert_legend(flanders_gardens, "BAF1_50", "A. Garden Adaptation Contribution",
                                      breaks_labels[["BAF1_50"]]$breaks,
                                      breaks_labels[["BAF1_50"]]$labels,
                                      show_legend = FALSE, show_y = TRUE)

p2 <- prepare_metric_plot_vert_legend(flanders_gardens, "BAF2_50", "B. Garden Adaptation Potential",
                                      breaks_labels[["BAF2_50"]]$breaks,
                                      breaks_labels[["BAF2_50"]]$labels,
                                      show_legend = FALSE, show_y = FALSE)

p3 <- prepare_metric_plot_vert_legend(flanders_gardens, "BAFper_50", "C. Garden Contribution Efficiency",
                                      breaks_labels[["BAFper_50"]]$breaks,
                                      breaks_labels[["BAFper_50"]]$labels,
                                      show_legend = FALSE, show_y = TRUE)

p4 <- prepare_metric_plot_vert_legend(flanders_gardens, "BAFtot_50", "D. Garden Potential Gap",
                                      breaks_labels[["BAFtot_50"]]$breaks,
                                      breaks_labels[["BAFtot_50"]]$labels,
                                      show_legend = TRUE, show_y = FALSE)  # Only this shows legend

first_row <- (p1 | p2)/(p3 | p4) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))
```

```{r}
Manchester_gardens <- Manchester_gardens %>%
  mutate(dataset = "Manchester") 

# --- Step 2: Plotting function ---
make_blue_palette <- function(n) {
  colorRampPalette(c("#c6d9e4", "#407ca2"))(n)  # light to dark red
}

base_size <- 30
# --- Function to create plots with horizontal legend (same as before, adapted) ---
prepare_metric_plot_vert_legend <- function(df, metric_col, metric_name, breaks, labels, show_legend = TRUE, show_y = FALSE) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- make_blue_palette(n_cats)
  
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 7, color = "black") +
    labs(x = NULL, y = if(show_y) "% of gardens" else NULL, title = metric_name) +
    scale_fill_manual(values = my_palette,
                      name = "Levels",
                      guide = if(show_legend) guide_legend(direction = "horizontal", nrow = 1) else "none") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = base_size-2, face = "bold"),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size),
      axis.text.x = element_text(angle = 30, hjust = 1, size = base_size),
      axis.title.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.text.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.ticks.y = if(show_y) element_line() else element_blank(),
      legend.text = element_text(size = base_size - 4),
      legend.title = element_text(size = base_size -2 ),
      legend.position = if(show_legend) "bottom" else "none"
    )
  
  return(p)
}
# --- Step 3: Define breaks/labels ---
breaks_labels <- list(
  BAF1_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAF2_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAFtot_50 = list(
    breaks = c(-Inf, 0.025, 0.05, 0.075, 0.1, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  ),
  BAFper_50 = list(
    breaks = c(-Inf, 20, 40, 60, 80, +Inf),
    labels = c("Very low", "Low", "Moderate", "High", "Very High")
  )
)

# Supply plots
p1 <- prepare_metric_plot_vert_legend(Manchester_gardens, "BAF1_50", "A. Garden Adaptation Contribution",
                                      breaks_labels[["BAF1_50"]]$breaks,
                                      breaks_labels[["BAF1_50"]]$labels,
                                      show_legend = FALSE, show_y = TRUE)

p2 <- prepare_metric_plot_vert_legend(Manchester_gardens, "BAF2_50", "B. Garden Adaptation Potential",
                                      breaks_labels[["BAF2_50"]]$breaks,
                                      breaks_labels[["BAF2_50"]]$labels,
                                      show_legend = FALSE, show_y = FALSE)

p3 <- prepare_metric_plot_vert_legend(Manchester_gardens, "BAFper_50", "C. Garden Contribution Efficiency",
                                      breaks_labels[["BAFper_50"]]$breaks,
                                      breaks_labels[["BAFper_50"]]$labels,
                                      show_legend = FALSE, show_y = TRUE)

p4 <- prepare_metric_plot_vert_legend(Manchester_gardens, "BAFtot_50", "D. Garden Potential Gap",
                                      breaks_labels[["BAFtot_50"]]$breaks,
                                      breaks_labels[["BAFtot_50"]]$labels,
                                      show_legend = TRUE, show_y = FALSE)  # Only this shows legend

first_row <- (p1 | p2)/(p3 | p4) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))
```





```{r}
# --- Custom red palette function ---
make_red_palette <- function(n) {
  colorRampPalette(c("#f7d9d9", "#c85a5a"))(n)  # light to dark red
}

# --- Function to create plots with horizontal legend (same as before, adapted) ---
prepare_metric_plot_red <- function(df, metric_col, metric_name, breaks, labels, show_legend = TRUE, show_y = FALSE) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- make_red_palette(n_cats)
  
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 5, color = "black") +
    labs(x = NULL, y = if(show_y) "% of gardens" else NULL, title = metric_name) +
    scale_fill_manual(values = my_palette,
                      name = "Levels",
                      guide = if(show_legend) guide_legend(direction = "horizontal", nrow = 1) else "none") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = base_size-2, face = "bold"),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size),
      axis.text.x = element_text(angle = 30, hjust = 1, size = base_size),
      axis.title.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.text.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.ticks.y = if(show_y) element_line() else element_blank(),
      legend.text = element_text(size = base_size - 4),
      legend.title = element_text(size = base_size -2 ),
      legend.position = if(show_legend) "bottom" else "none"
    )
  
  return(p)
}

library(patchwork)

new_breaks_labels <- list(
  vul50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1")
  ),
  exp50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1")
  ),
  LSTrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1")
  ),
  TWIrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1")
  )
)

# Only first plot shows y-axis
r1 <- prepare_metric_plot_red(combined_data, "LSTrisk50", "E. Heat risk",
                              new_breaks_labels[["LSTrisk50"]]$breaks,
                              new_breaks_labels[["LSTrisk50"]]$labels,
                              show_legend = FALSE, show_y = TRUE)

r2 <- prepare_metric_plot_red(combined_data, "TWIrisk50", "F. Flood Risk",
                              new_breaks_labels[["TWIrisk50"]]$breaks,
                              new_breaks_labels[["TWIrisk50"]]$labels,
                              show_legend = FALSE, show_y = FALSE)

r3 <- prepare_metric_plot_red(combined_data, "vul50", "G. Social Vulnerability",
                              new_breaks_labels[["vul50"]]$breaks,
                              new_breaks_labels[["vul50"]]$labels,
                              show_legend = FALSE, show_y = FALSE)

r4 <- prepare_metric_plot_red(combined_data, "exp50", "H. Exposure",
                              new_breaks_labels[["exp50"]]$breaks,
                              new_breaks_labels[["exp50"]]$labels,
                              show_legend = TRUE, show_y = FALSE)

# Combine horizontally with one common legend
second_row <- r1 | r2 | r3 | r4
second_row <- second_row +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))

# Combine with first row
combined_plot <- (p1 | p2 | p3 | p4) / second_row + plot_layout(heights = c(1, 1))

combined_plot

ggsave("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Figures/SupplyDemand.png",
       plot = combined_plot, width = 20, height = 12, dpi = 300)

sum(is.na(all_gardens_15$exp50))

```

```{r}
# Mapping of NISCODE to city names
niscode_names <- c(
  "11002" = "Antwerp",
  "12025" = "Mechelen",
  "13040" = "Turnhout",
  "24062" = "Leuven",
  "31005" = "Bruges",
  "34022" = "Kortrijk",
  "35013" = "Ostend",
  "36015" = "Roeselare",
  "41002" = "Aalst",
  "44021" = "Ghent",
  "46021" = "Sint-Niklaas",
  "71016" = "Genk",
  "71022" = "Hasselt"
)

# Filter and map
flanders_gardens <- st_drop_geometry(all_gardens_15) %>%
  filter(NISCODE %in% names(niscode_names)) %>%        # keep only known cities
  mutate(dataset = niscode_names[as.character(NISCODE)])  # replace NISCODE with city names

# --- Custom red palette function ---
make_red_palette <- function(n) {
  colorRampPalette(c("#f7d9d9", "#c85a5a"))(n)  # light to dark red
}

base_size <- 26
# --- Function to create plots with horizontal legend (same as before, adapted) ---
prepare_metric_plot_red <- function(df, metric_col, metric_name, breaks, labels, show_legend = TRUE, show_y = FALSE) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- make_red_palette(n_cats)
  
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 5, color = "black") +
    labs(x = NULL, y = if(show_y) "% of gardens" else NULL, title = metric_name) +
    scale_fill_manual(values = my_palette,
                      name = "Levels",
                      guide = if(show_legend) guide_legend(direction = "horizontal", nrow = 1) else "none") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = base_size-2, face = "bold"),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size),
      axis.text.x = element_text(angle = 30, hjust = 1, size = base_size),
      axis.title.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.text.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.ticks.y = if(show_y) element_line() else element_blank(),
      legend.text = element_text(size = base_size - 4),
      legend.title = element_text(size = base_size -2 ),
      legend.position = if(show_legend) "bottom" else "none"
    )
  
  return(p)
}

library(patchwork)

new_breaks_labels <- list(
  vul50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  exp50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  LSTrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  TWIrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  )
)

# Demand plots
r1 <- prepare_metric_plot_red(flanders_gardens, "LSTrisk50", "E. Heat Risk",
                              new_breaks_labels[["LSTrisk50"]]$breaks,
                              new_breaks_labels[["LSTrisk50"]]$labels,
                              show_legend = FALSE, show_y = TRUE)

r2 <- prepare_metric_plot_red(flanders_gardens, "TWIrisk50", "F. Flood Risk",
                              new_breaks_labels[["TWIrisk50"]]$breaks,
                              new_breaks_labels[["TWIrisk50"]]$labels,
                              show_legend = FALSE, show_y = FALSE)

r3 <- prepare_metric_plot_red(flanders_gardens, "vul50", "G. Social Vulnerability",
                              new_breaks_labels[["vul50"]]$breaks,
                              new_breaks_labels[["vul50"]]$labels,
                              show_legend = FALSE, show_y = TRUE)

r4 <- prepare_metric_plot_red(flanders_gardens, "exp50", "H. Exposure",
                              new_breaks_labels[["exp50"]]$breaks,
                              new_breaks_labels[["exp50"]]$labels,
                              show_legend = TRUE, show_y = FALSE)  # Only this shows legend

second_row <- (r1 | r2)/(r3 | r4) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))

library(patchwork)

# --- Supply row (2x2) ---
library(patchwork)

supply_block <- ((p1 | p2) / (p3 | p4)) +
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(
      legend.position = "bottom"
    )
  )

# --- Demand row (2x2) ---
demand_block <- ((r1 | r2) / (r3 | r4)) +
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(
      legend.position = "bottom"
    )
  )


# --- Combine rows vertically ---
combined_plot <- wrap_plots(supply_block, demand_block, ncol = 1, nrow = 2)

# --- Save ---
ggsave("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Figures/results/SupplyDemand_Flanders.png",
       plot = combined_plot,
       width = 22, height = 28, dpi = 300)

```

```{r}

Manchester_gardens <- Manchester_gardens %>%
  mutate(dataset = "Manchester") 

# --- Custom red palette function ---
make_red_palette <- function(n) {
  colorRampPalette(c("#f7d9d9", "#c85a5a"))(n)  # light to dark red
}

base_size <- 30
# --- Function to create plots with horizontal legend (same as before, adapted) ---
prepare_metric_plot_red <- function(df, metric_col, metric_name, breaks, labels, show_legend = TRUE, show_y = FALSE) {
  
  df_long <- df %>%
    select(dataset, all_of(metric_col)) %>%
    pivot_longer(cols = all_of(metric_col),
                 names_to = "metric",
                 values_to = "value") %>%
    mutate(metric = metric_name)
  
  df_long <- df_long %>%
    mutate(
      BAF_cat = cut(value,
                    breaks = breaks,
                    labels = labels,
                    include.lowest = TRUE,
                    dig.lab = 3)
    )
  
  cat_perc <- df_long %>%
    group_by(dataset, metric, BAF_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(dataset, metric) %>%
    mutate(perc = 100 * n / sum(n)) %>%
    ungroup()
  
  n_cats <- length(unique(cat_perc$BAF_cat))
  my_palette <- make_red_palette(n_cats)
  
  p <- ggplot(cat_perc, aes(x = dataset, y = perc, fill = BAF_cat)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = ifelse(perc > 2, sprintf("%.1f", perc), "")),
              position = position_stack(vjust = 0.5),
              size = 7, color = "black") +
    labs(x = NULL, y = if(show_y) "% of gardens" else NULL, title = metric_name) +
    scale_fill_manual(values = my_palette,
                      name = "Levels",
                      guide = if(show_legend) guide_legend(direction = "horizontal", nrow = 1) else "none") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = base_size-2, face = "bold"),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size),
      axis.text.x = element_text(angle = 30, hjust = 1, size = base_size),
      axis.title.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.text.y = if(show_y) element_text(size = base_size) else element_blank(),
      axis.ticks.y = if(show_y) element_line() else element_blank(),
      legend.text = element_text(size = base_size - 4),
      legend.title = element_text(size = base_size -2 ),
      legend.position = if(show_legend) "bottom" else "none"
    )
  
  return(p)
}

library(patchwork)

new_breaks_labels <- list(
  vul50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  exp50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  LSTrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  ),
  TWIrisk50 = list(
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, +Inf),
    labels = c("Very Low", "Low", "Moderate", "High", "Very High")
  )
)

# Demand plots
r1 <- prepare_metric_plot_red(Manchester_gardens, "LSTrisk50", "E. Heat Risk",
                              new_breaks_labels[["LSTrisk50"]]$breaks,
                              new_breaks_labels[["LSTrisk50"]]$labels,
                              show_legend = FALSE, show_y = TRUE)

r2 <- prepare_metric_plot_red(Manchester_gardens, "TWIrisk50", "F. Flood Risk",
                              new_breaks_labels[["TWIrisk50"]]$breaks,
                              new_breaks_labels[["TWIrisk50"]]$labels,
                              show_legend = FALSE, show_y = FALSE)

r3 <- prepare_metric_plot_red(Manchester_gardens, "vul50", "G. Social Vulnerability",
                              new_breaks_labels[["vul50"]]$breaks,
                              new_breaks_labels[["vul50"]]$labels,
                              show_legend = FALSE, show_y = TRUE)

r4 <- prepare_metric_plot_red(Manchester_gardens, "exp50", "H. Exposure",
                              new_breaks_labels[["exp50"]]$breaks,
                              new_breaks_labels[["exp50"]]$labels,
                              show_legend = TRUE, show_y = FALSE)  # Only this shows legend

second_row <- (r1 | r2)/(r3 | r4) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))

library(patchwork)

supply_block <- ((p1 | p2) / (p3 | p4)) +
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(
      legend.position = "bottom"
    )
  )

# --- Demand row (2x2) ---
demand_block <- ((r1 | r2) / (r3 | r4)) +
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(
      legend.position = "bottom"
    )
  )


# --- Combine rows vertically ---
combined_plot <- wrap_plots(supply_block, demand_block, ncol = 1, nrow = 2)

# --- Save ---
ggsave("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Figures/SupplyDemand_Manchester.png",
       plot = combined_plot,
       width = 22, height = 28, dpi = 300)

```


#STEP 3: Tabel of % HG, % water, ...

```{r}
library(dplyr)
library(tidyr)
library(stringr)

# Drop geometry and clean Flanders city names
flanders_df <- st_drop_geometry(all_gardens_15) %>%
  mutate(
    region = "Flanders",
    city = str_replace(city, "ADPgrdns$", "")
  )

manchester_df <- st_drop_geometry(Manchester_gardens) %>%
  mutate(region = "Manchester")

# Combine attribute tables
combined_data <- bind_rows(flanders_df, manchester_df)

# Calculate % of each cover type per garden
combined_data <- combined_data %>%
  mutate(
    perc_HG = HG50 / ex50 * 100,
    perc_LG = LG50 / ex50 * 100,
    perc_W  = W50  / ex50 * 100,
    perc_BS = BS50 / ex50 * 100,
    perc_A = A50/ex50 *100
  )

# Make 'region' a factor to control the order
combined_data <- combined_data %>%
  mutate(region = factor(region, levels = c("Flanders", "Manchester")))

# Summarise mean and SD per city
summary_stats <- combined_data %>%
  group_by(city, region) %>%
  summarise(
    `Garden Area Mean` = mean(ex50),
    `Garden Area SD`   = sd(ex50),
    `% High Green Mean` = mean(perc_HG),
    `% High Green SD`   = sd(perc_HG),
    `% Low Green Mean`  = mean(perc_LG),
    `% Low Green SD`    = sd(perc_LG),
    `% Water Mean`      = mean(perc_W),
    `% Water SD`        = sd(perc_W),
    `% Bare Soil Mean`  = mean(perc_BS),
    `% Bare Soil SD`    = sd(perc_BS),
    `% Sealed Surface Mean`  = mean(perc_A),
    `% Sealed Surface SD`    = sd(perc_A),
    .groups = "drop"
  ) %>%
  arrange(region, city)  # ensures Flanders first, Manchester last

# View results
summary_stats

library(writexl)

# Save summary_stats to Excel
write_xlsx(summary_stats, "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Tabellen/summary_stats.xlsx")

```

#Heatmap of contingency table

```{r}
# Load necessary library
library(ggplot2)
library(dplyr)

# Supply classification
classify_supply <- function(x) {
  cut(
    x,
    breaks = c(0, 0.025, 0.05, 0.075, Inf),  # custom breaks for supply
    include.lowest = TRUE,
    labels = c("0-0.025", "0.025-0.05", "0.075-0.1", ">0.1")
  )
}

# Demand classification
classify_demand <- function(x) {
  cut(
    x,
    breaks = c(0, 0.25, 0.5, 0.75, 1),  # custom breaks for demand
    include.lowest = TRUE,
    labels = c("0-0.25", "0.25-0.5", "0.5-0.75", "0.75-1")
  )
}

# Apply to datasets
all_gardens_15 <- all_gardens_15 %>%
  mutate(
    supply_class = classify_supply(BAF1_50),
    demand_class = classify_demand(risk50)
  )

Manchester_gardens <- Manchester_gardens %>%
  mutate(
    supply_class = classify_supply(BAF1_50),
    demand_class = classify_demand(risk50)
  )


# Function to plot heatmap with numbers
plot_heatmap <- function(data, title="Heatmap") {
  data_summary <- data %>%
    count(supply_class, demand_class)
  
  ggplot(data_summary, aes(x = supply_class, y = demand_class, fill = n)) +
    geom_tile(color = "white") +
    geom_text(aes(label = n), color = "black", size = 5) +  # Add numbers
    scale_fill_gradient(low = "white", high = "#8b5fbf") +
    labs(
      x = "Supply (Garden Contribution)",
      y = "Demand (CVI)",
      fill = "Count",
      title = title
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Plot for each dataset
plot_heatmap(all_gardens_15, "Flemish cities")
plot_heatmap(Manchester_gardens, "Manchester")

```

```{r}
# Load necessary library
library(ggplot2)
library(dplyr)

# Supply classification
classify_supply <- function(x) {
  cut(
    x,
    breaks = c(0, 0.025, 0.05, 0.075, 1, Inf),  # custom breaks for supply
    include.lowest = TRUE,
    labels = c("0-0.025", "0.025-0.05","0.5-0.75","0.075-0.1", ">0.1")
  )
}

# Demand classification
classify_demand <- function(x) {
  cut(
    x,
    breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),  # custom breaks for demand
    include.lowest = TRUE,
    labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1")
  )
}

# Apply to datasets
all_gardens_15 <- all_gardens_15 %>%
  mutate(
    supply_class = classify_supply(BAF1_50),
    demand_class = classify_demand(risk50)
  )

Manchester_gardens <- Manchester_gardens %>%
  mutate(
    supply_class = classify_supply(BAF1_50),
    demand_class = classify_demand(risk50)
  )


# Function to plot heatmap with numbers
plot_heatmap <- function(data, title="Heatmap") {
  data_summary <- data %>%
    count(supply_class, demand_class)
  
  ggplot(data_summary, aes(x = supply_class, y = demand_class, fill = n)) +
    geom_tile(color = "white") +
    geom_text(aes(label = n), color = "black", size = 5) +  # Add numbers
    scale_fill_gradient(low = "white", high = "#8b5fbf") +
    labs(
      x = "Supply (Garden Contribution)",
      y = "Demand (CVI)",
      fill = "Count",
      title = title
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Plot for each dataset
plot_heatmap(all_gardens_15, "Flemish cities")
plot_heatmap(Manchester_gardens, "Manchester")
```

```{r}
library(ggplot2)
library(dplyr)
library(patchwork)

# Supply classification
classify_supply <- function(x) {
  cut(
    x,
    breaks = c(0, 0.025, 0.05, 0.075, 0.1, Inf),  # custom breaks for supply
    include.lowest = TRUE,
    labels = c("Very Low", "Low","Moderate","High", "Very High")
  )
}

# Demand classification
classify_demand <- function(x) {
  cut(
    x,
    breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),  # custom breaks for demand
    include.lowest = TRUE,
    labels = c("Very Low", "Low","Moderate","High", "Very High")
  )
}


# Function to compute percentages
compute_percent <- function(data, supply_col, demand_col) {
  data %>%
    mutate(
      supply_class = classify_supply(.data[[supply_col]]),
      demand_class = classify_demand(.data[[demand_col]])
    ) %>%
    count(supply_class, demand_class) %>%
    mutate(percent = n / sum(n) * 100)
}

# Compute data summaries
df1 <- compute_percent(all_gardens_15, "BAF1_50", "risk50")
df2 <- compute_percent(Manchester_gardens, "BAF1_50", "risk50")
df3 <- compute_percent(all_gardens_15, "BAF2_50", "risk50")
df4 <- compute_percent(Manchester_gardens, "BAF2_50", "risk50")

# Find max percent for consistent color scale
max_percent <- max(c(df1$percent, df2$percent, df3$percent, df4$percent))

# Function to plot heatmap with fixed fill scale
plot_heatmap_fixed <- function(df, supply_label, title) {
  ggplot(df, aes(x = supply_class, y = demand_class, fill = percent)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.1f%%", percent)), color = "black", size = 4) +
    scale_fill_gradient(low = "white", high = "#855a6e", limits = c(0, max_percent)) +
    labs(
      x = supply_label,
      y = "Demand (CVI)",
      fill = "Percent",
      title = title
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Create the four heatmaps
heat1 <- plot_heatmap_fixed(df1, "Supply (Garden Adaptation Contribution)", "A. Flanders Current")
heat2 <- plot_heatmap_fixed(df2, "Supply (Garden Adaptation Contribution)", "B. Manchester Current")
heat3 <- plot_heatmap_fixed(df3, "Supply (Garden Adaptation Potential)", "C. Flanders Potential")
heat4 <- plot_heatmap_fixed(df4, "Supply (Garden Adaptation Potential)", "D. Manchester Potential")

# Combine into 2x2 figure with a single shared legend
# Apply theme to each plot manually
plots <- list(heat1, heat2, heat3, heat4) |>
  lapply(\(p) p + theme(legend.position = "right"))

combined_heatmaps <- (
  (plots[[1]] | plots[[2]]) /
  (plots[[3]] | plots[[4]])
) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "right"))

# Display
combined_heatmaps

# Create the two heatmaps
heat1 <- plot_heatmap_fixed(df1, "Supply (Garden Adaptation Contribution)", "A. Flanders Current")
heat3 <- plot_heatmap_fixed(df3, "Supply (Garden Adaptation Potential)", "B. Flanders Potential")

# Combine into 2x2 figure with a single shared legend
# Apply theme to each plot manually
plots <- list(heat1, heat3) |>
  lapply(\(p) p + theme(legend.position = "right"))

combined_heatmaps <- (
  (plots[[1]] | plots[[2]]) 
) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "right"))

# Display
combined_heatmaps

```


```{r}
library(ggplot2)
library(dplyr)
library(patchwork)

# Supply classification
classify_supply <- function(x) {
  cut(
    x,
    breaks = c(0, 0.025, 0.05, 0.075, 0.1, Inf),  # custom breaks for supply
    include.lowest = TRUE,
    labels = c("Very Low", "Low","Moderate","High", "Very High")
  )
}

# Demand classification
classify_demand <- function(x) {
  cut(
    x,
    breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),  # custom breaks for demand
    include.lowest = TRUE,
    labels = c("Very Low", "Low","Moderate","High", "Very High")
  )
}


# Function to compute percentages
compute_percent <- function(data, supply_col, demand_col) {
  data %>%
    mutate(
      supply_class = classify_supply(.data[[supply_col]]),
      demand_class = classify_demand(.data[[demand_col]])
    ) %>%
    count(supply_class, demand_class) %>%
    mutate(percent = n / sum(n) * 100)
}

# Compute data summaries
df1 <- compute_percent(all_gardens_15, "BAF1_50", "risk50")
df2 <- compute_percent(Manchester_gardens, "BAF1_50", "risk50")
df3 <- compute_percent(all_gardens_15, "BAF2_50", "risk50")
df4 <- compute_percent(Manchester_gardens, "BAF2_50", "risk50")

# Find max percent for consistent color scale
max_percent <- max(c(df1$percent, df2$percent, df3$percent, df4$percent))

# Function to plot heatmap with fixed fill scale
plot_heatmap_fixed <- function(df, supply_label, title) {
  ggplot(df, aes(x = supply_class, y = demand_class, fill = percent)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.1f%%", percent)), color = "black", size = 4) +
    scale_fill_gradient(low = "white", high = "#855a6e", limits = c(0, max_percent)) +
    labs(
      x = supply_label,
      y = "Demand (CVI)",
      fill = "Percent",
      title = title
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Create the four heatmaps
heat1 <- plot_heatmap_fixed(df1, "Supply (Garden Adaptation Contribution)", "A. Flanders Current")
heat2 <- plot_heatmap_fixed(df2, "Supply (Garden Adaptation Contribution)", "B. Manchester Current")
heat3 <- plot_heatmap_fixed(df3, "Supply (Garden Adaptation Potential)", "C. Flanders Potential")
heat4 <- plot_heatmap_fixed(df4, "Supply (Garden Adaptation Potential)", "D. Manchester Potential")

# Combine into 2x2 figure with a single shared legend
# Apply theme to each plot manually
plots <- list(heat1, heat2, heat3, heat4) |>
  lapply(\(p) p + theme(legend.position = "right"))

combined_heatmaps <- (
  (plots[[1]] | plots[[2]]) /
  (plots[[3]] | plots[[4]])
) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "right"))

# Display
combined_heatmaps

# Create the two heatmaps
heat2 <- plot_heatmap_fixed(df2, "Supply (Garden Adaptation Contribution)", "A. Manchester Current")
heat4 <- plot_heatmap_fixed(df4, "Supply (Garden Adaptation Potential)", "B. Manchester Potential")

# Combine into 2x2 figure with a single shared legend
# Apply theme to each plot manually
plots <- list(heat2, heat4) |>
  lapply(\(p) p + theme(legend.position = "right"))

combined_heatmaps <- (
  (plots[[1]] | plots[[2]]) 
) +
  plot_layout(guides = "collect") +
  plot_annotation(theme = theme(legend.position = "right"))

# Display
combined_heatmaps
```


#Spearman rank correlation

```{r}
#1
gardens_F_analysis <-all_gardens_15[,c("BAF1_50","BAF2_50","BAFper_50","BAFtot_50","TWIrisk50","LSTrisk50","vul50","exp50")]

# Rename columns
gardens_F_analysis <- st_drop_geometry(gardens_F_analysis)
colnames(gardens_F_analysis) <- c(
  "Garden Adaptation Contribution",
  "Garden Adaptation Potential",
  "Garden Contribution Efficiency",
  "Garden Potential Gap",
  "Flood risk",
  "Heat risk",
  "Social vulnerability",
  "Exposure"
)
# Convert all columns to numeric
gardens_F_analysis[] <- lapply(gardens_F_analysis, function(x) as.numeric(as.character(x)))

# Calculate Spearman correlation coefficients
cor_matrix <- cor(gardens_F_analysis, method = "spearman", use="complete.obs")

# Open PNG device
png("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Visualisations/corrplot_matrix.png", width = 1000, height = 800, res = 150)


cor <- corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  tl.cex = 0.8,       # size of axis labels
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black", 
  number.cex = 0.6    # size of numbers inside cells
)

cor

# Close device
dev.off()
```

```{r}
library(dplyr)
library(corrplot)
library(sf)

# Assuming your dataset has a column 'NISCODE'
gardens_F_analysis <- all_gardens_15[, c(
  "NISCODE",
  "BAF1_50","BAF2_50","BAFper_50","BAFtot_50",
  "TWIrisk50","LSTrisk50","vul50","exp50"
)]

# Remove geometry
gardens_F_analysis <- st_drop_geometry(gardens_F_analysis)

# Rename columns
colnames(gardens_F_analysis)[2:9] <- c(
  "Garden Adaptation Contribution",
  "Garden Adaptation Potential",
  "Garden Contribution Efficiency",
  "Garden Potential Gap",
  "Flood risk",
  "Heat risk",
  "Social vulnerability",
  "Exposure"
)

# Convert all to numeric
gardens_F_analysis[2:9] <- lapply(gardens_F_analysis[2:9], function(x) as.numeric(as.character(x)))

# Create output folder
output_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Visualisations/corrplots_per_NISCODE/"
dir.create(output_dir, showWarnings = FALSE)

# Loop over each NISCODE
unique_codes <- unique(gardens_F_analysis$NISCODE)

for (code in unique_codes) {
  subset_data <- gardens_F_analysis %>% filter(NISCODE == code) %>% select(-NISCODE)
  
  # Only proceed if there are at least 2 rows (corrplot needs >1 row)
  if (nrow(subset_data) > 1) {
    
    cor_matrix <- cor(subset_data, method = "spearman", use = "complete.obs")
    
    # Define file name
    file_name <- paste0(output_dir, "corrplot_NISCODE_", code, ".png")
    
    # Open PNG device
    png(file_name, width = 1000, height = 800, res = 150)
    
    # Plot correlation matrix
    corrplot(
      cor_matrix,
      method = "color",
      type = "upper",
      tl.cex = 0.8,
      tl.col = "black",
      tl.srt = 45,
      addCoef.col = "black",
      number.cex = 0.6
    )
    
    dev.off()
    
    message("Saved correlation plot for NISCODE: ", code)
  } else {
    message("Skipping NISCODE ", code, " (not enough data)")
  }
}

```


```{r}
#1
gardens_F_analysis <-Manchester_gardens[,c("BAF1_50","BAF2_50","BAFtot_50","BAFper_50","TWIrisk50","LSTrisk50","vul50", "exp50")]

# Rename columns
gardens_F_analysis <- st_drop_geometry(gardens_F_analysis)
colnames(gardens_F_analysis) <- c(
  "Garden Adaptation Contribution",
  "Garden Adaptation Potential",
  "Garden Contribution Efficiency",
  "Garden Potential Gap",
  "Flood risk",
  "Heat risk",
  "Social vulnerability",
  "Exposure"
)
# Convert all columns to numeric
gardens_F_analysis[] <- lapply(gardens_F_analysis, function(x) as.numeric(as.character(x)))

# Calculate Spearman correlation coefficients
cor_matrix <- cor(gardens_F_analysis, method = "spearman", use="complete.obs")

# Open PNG device
png("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Visualisations/corrplot_matrix_M.png", width = 1000, height = 800, res = 150)

cor <- corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  tl.cex = 0.8,       # size of axis labels
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black", 
  number.cex = 0.6    # size of numbers inside cells
)

cor

# Close device
dev.off()
```

```{r}
library(corrplot)

# Prepare data for Flanders
gardens_F_analysis <- all_gardens_15[,c("BAF1_50","BAF2_50","BAFtot_50","BAFper_50","TWIrisk50","LSTrisk50","vul50","exp50")]
gardens_F_analysis <- st_drop_geometry(gardens_F_analysis)
colnames(gardens_F_analysis) <- c(
  "Garden Contribution",
  "Garden Potential",
  "PGI",
  "CEI",
  "Flood risk",
  "Heat risk",
  "Social vulnerability",
  "Exposure"
)
gardens_F_analysis[] <- lapply(gardens_F_analysis, function(x) as.numeric(as.character(x)))
cor_matrix_FL <- cor(gardens_F_analysis, method = "spearman", use="complete.obs")

# Prepare data for Manchester
gardens_M_analysis <- Manchester_gardens[,c("BAF1_50","BAF2_50","BAFtot_50","BAFper_50","TWIrisk50","LSTrisk50","vul50", "exp50")]
gardens_M_analysis <- st_drop_geometry(gardens_M_analysis)
colnames(gardens_M_analysis) <- c(
  "Garden Contribution",
  "Garden Potential",
  "PGI",
  "CEI",
  "Flood risk",
  "Heat risk",
  "Social vulnerability",
  "Exposure"
)
gardens_M_analysis[] <- lapply(gardens_M_analysis, function(x) as.numeric(as.character(x)))
cor_matrix_M <- cor(gardens_M_analysis, method = "spearman", use="complete.obs")

# Determine common color scale limits
min_corr <- min(c(cor_matrix_FL, cor_matrix_M))
max_corr <- max(c(cor_matrix_FL, cor_matrix_M))

# Plot side by side with shared color scale
par(mfrow=c(2,1), mar=c(2,2,2,2))  # 1 row, 2 columns

corrplot(
  cor_matrix_FL,
  method = "color",
  type = "upper",
  tl.cex = 0.8,
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black", 
  number.cex = 0.6,
  cl.lim = c(min_corr, max_corr),
  title = "A. Flanders"
)

corrplot(
  cor_matrix_M,
  method = "color",
  type = "upper",
  tl.cex = 0.8,
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black", 
  number.cex = 0.6,
  cl.lim = c(min_corr, max_corr),
  title = "B. Manchester"
)

```

```{r}
summary_stats_FL <- all_gardens_15 %>%
  group_by(city) %>%
  summarise(
    mean_UGEI_50 = mean(UGEI_50h, na.rm=TRUE),
    sd_UGEI_50 = sd(UGEI_50h, na.rm=TRUE),
    mean_BAF1_50  = mean(BAF1_50, na.rm = TRUE),
    sd_BAF1_50    = sd(BAF1_50, na.rm = TRUE),
    mean_BAF2_50  = mean(BAF2_50, na.rm = TRUE),
    sd_BAF2_50    = sd(BAF2_50, na.rm = TRUE),
    mean_risk50   = mean(risk50, na.rm = TRUE),
    sd_risk50     = sd(risk50, na.rm = TRUE)
  )

summary_stats_FL_2 <- all_gardens_15 %>%
  summarise(
    mean_UGEI_50 = mean(UGEI_50h, na.rm=TRUE),
    sd_UGEI_50 = sd(UGEI_50h, na.rm=TRUE),
    mean_BAF1_50  = mean(BAF1_50, na.rm = TRUE),
    sd_BAF1_50    = sd(BAF1_50, na.rm = TRUE),
    mean_BAF2_50  = mean(BAF2_50, na.rm = TRUE),
    sd_BAF2_50    = sd(BAF2_50, na.rm = TRUE),
    mean_risk50   = mean(risk50, na.rm = TRUE),
    sd_risk50     = sd(risk50, na.rm = TRUE)
  )
summary_stats_FL_2$city <- "Flemish cities"

summary_stats_M <- Manchester_gardens %>%
  group_by(city) %>%
  summarise(
    mean_UGEI_50 = mean(UGEI_50h, na.rm=TRUE),
    sd_UGEI_50 = sd(UGEI_50h, na.rm=TRUE),
    mean_BAF1_50  = mean(BAF1_50, na.rm = TRUE),
    sd_BAF1_50    = sd(BAF1_50, na.rm = TRUE),
    mean_BAF2_50  = mean(BAF2_50, na.rm = TRUE),
    sd_BAF2_50    = sd(BAF2_50, na.rm = TRUE),
    mean_risk50   = mean(risk50, na.rm = TRUE),
    sd_risk50     = sd(risk50, na.rm = TRUE)
  )

summary_stats_FL <- st_drop_geometry(summary_stats_FL)
summary_stats_FL_2 <- st_drop_geometry(summary_stats_FL_2)
summary_stats_M <- st_drop_geometry(summary_stats_M)

summary_stats <- rbind(summary_stats_FL, summary_stats_M)
summary_stats_2 <- rbind(summary_stats_FL_2, summary_stats_M)
```


```{r}
# drop geometry if it's still an sf object
summary_table <- st_drop_geometry(summary_stats)

# write to Excel
write_xlsx(summary_table, "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/summary_stats.xlsx")
```

```{r}
# drop geometry before pivoting
summary_long <- summary_stats %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value)


# Create a named vector mapping old city codes/names to nicer labels
city_labels <- c(
  AalstADPgrdns = "Aalst",
  AntwerpenADPgrdns = "Antwerpen",
  BruggeADPgrdns = "Brugge",
  GenkADPgrdns ="Genk",
  GentADPgrdns ="Gent",
  HasseltADPgrdns ="Hasselt",
  KortrijkADPgrdns ="Kortrijk",
  LeuvenADPgrdns ="Leuven",
  MechelenADPgrdns="Mechelen",
  OostendeADPgrdns ="Oostende",
  RoeselareADPgrdns ="Roeselare",
  Sint = "Sint-Niklaas",
  TurnhoutADPgrdns ="Turnhout",
  Manchester ="Manchester"
  # add all your cities here
)

# Define nicer variable labels
variable_labels <- c(
  "BAFper_50" = "Supply: Contribution efficiency index",
  "BAFtot_50" = "Supply: Realization Gap Index",
  "risk50"  = "Demand",
  "UGEI_50"      = "Equity Index"
)

ggplot(summary_long, aes(x = city, y = mean, fill = variable)) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  facet_wrap(~ variable, scales = "free_y",
             labeller = as_labeller(variable_labels)) +
  labs(y = "Mean ± SD", x = "City", title = "Garden Metrics per City") +
  scale_fill_manual(
    values = c(
      "BAFper_50" = "#407ca2",
      "BAFtot_50" = "#6aa6c8",
      "risk50"  = "#bc3e45",
      "UGEI_50"      = "#855a6e"
    )
  ) +
  scale_x_discrete(labels = city_labels) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"   # << hides legend
  )
```
```{r}
# Prepare long data
summary_long2 <- summary_stats %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    # Flip demand negative
    mean = ifelse(variable == "risk50", -mean, mean)
  )

# Add spacer rows
var_levels <- summary_long2 %>% distinct(variable)
spacers <- var_levels %>% mutate(city = "Spacer", mean = NA_real_, sd = NA_real_)
summary_long2 <- bind_rows(summary_long2, spacers)

# Combine Supply & Demand into one group
summary_combined <- summary_long2 %>%
  filter(variable %in% c("BAF1_50", "risk50", "UGEI_50")) %>%
  mutate(variable_group = ifelse(variable %in% c("BAF1_50", "risk50"),
                                 "SupplyDemand", variable))

# City ordering
city_levels <- c(
  "AalstADPgrdns","AntwerpenADPgrdns","BruggeADPgrdns","GenkADPgrdns","GentADPgrdns",
  "HasseltADPgrdns","KortrijkADPgrdns","LeuvenADPgrdns","MechelenADPgrdns","OostendeADPgrdns",
  "RoeselareADPgrdns","Sint","TurnhoutADPgrdns",
  "Spacer",
  "Manchester"
)

# City labels
city_labels <- c(
  AalstADPgrdns = "Aalst",
  AntwerpenADPgrdns = "Antwerpen",
  BruggeADPgrdns = "Brugge",
  GenkADPgrdns ="Genk",
  GentADPgrdns ="Gent",
  HasseltADPgrdns ="Hasselt",
  KortrijkADPgrdns ="Kortrijk",
  LeuvenADPgrdns ="Leuven",
  MechelenADPgrdns="Mechelen",
  OostendeADPgrdns ="Oostende",
  RoeselareADPgrdns ="Roeselare",
  Sint = "Sint-Niklaas",
  TurnhoutADPgrdns ="Turnhout",
  Spacer = " ",
  Manchester ="Manchester"
)

# --- Calculate symmetric y-axis for SupplyDemand ---
max_abs <- summary_combined %>%
  filter(variable %in% c("BAF1_50","risk50")) %>%
  summarise(max_val = max(abs(mean + sd), na.rm = TRUE)) %>%
  pull(max_val)

# Add a small padding for error bars
padding <- 0.7
max_abs_padded <- max_abs * (1 + padding)

a <- ggplot(summary_combined, aes(x = factor(city, levels = city_levels), y = mean, fill = variable)) +
  geom_col(position = "identity") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
  geom_hline(yintercept = 0, color = "black") +
  facet_wrap(~ variable_group, scales = "free_y",
             labeller = as_labeller(c(
               "SupplyDemand" = "Supply (+) vs Demand (–)",
               "UGEI_50"      = "Equity Index"
             ))) +
  scale_fill_manual(values = c(
    "BAF1_50" = "#407ca2",  # Supply
    "risk50"    = "#bc3e45",  # Demand
    "UGEI_50"   = "#855a6e"   # Equity
  )) +
  scale_x_discrete(labels = city_labels, drop = FALSE) +
  coord_cartesian(ylim = c(-max_abs_padded, max_abs_padded)) +
  labs(y = "Mean ± SD", x = "City", title = "A. Current Supply vs Demand and Equity Index") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```



```{r}
# Prepare long data
summary_long2 <- summary_stats %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    # Flip demand negative
    mean = ifelse(variable == "risk50", -mean, mean)
  )

# Add spacer rows
var_levels <- summary_long2 %>% distinct(variable)
spacers <- var_levels %>% mutate(city = "Spacer", mean = NA_real_, sd = NA_real_)
summary_long2 <- bind_rows(summary_long2, spacers)

# Combine Supply & Demand into one group
summary_combined <- summary_long2 %>%
  filter(variable %in% c("BAF2_50", "risk50", "UGEI_50")) %>%
  mutate(variable_group = ifelse(variable %in% c("BAF2_50", "risk50"),
                                 "SupplyDemand", variable))

# City ordering
city_levels <- c(
  "AalstADPgrdns","AntwerpenADPgrdns","BruggeADPgrdns","GenkADPgrdns","GentADPgrdns",
  "HasseltADPgrdns","KortrijkADPgrdns","LeuvenADPgrdns","MechelenADPgrdns","OostendeADPgrdns",
  "RoeselareADPgrdns","Sint","TurnhoutADPgrdns",
  "Spacer",
  "Manchester"
)

# City labels
city_labels <- c(
  AalstADPgrdns = "Aalst",
  AntwerpenADPgrdns = "Antwerpen",
  BruggeADPgrdns = "Brugge",
  GenkADPgrdns ="Genk",
  GentADPgrdns ="Gent",
  HasseltADPgrdns ="Hasselt",
  KortrijkADPgrdns ="Kortrijk",
  LeuvenADPgrdns ="Leuven",
  MechelenADPgrdns="Mechelen",
  OostendeADPgrdns ="Oostende",
  RoeselareADPgrdns ="Roeselare",
  Sint = "Sint-Niklaas",
  TurnhoutADPgrdns ="Turnhout",
  Spacer = " ",
  Manchester ="Manchester"
)

# --- Calculate symmetric y-axis for SupplyDemand ---
max_abs <- summary_combined %>%
  filter(variable %in% c("BAF2_50","risk50")) %>%
  summarise(max_val = max(abs(mean + sd), na.rm = TRUE)) %>%
  pull(max_val)

# Add a small padding for error bars
padding <- 0.7
max_abs_padded <- max_abs * (1 + padding)

b <- ggplot(summary_combined, aes(x = factor(city, levels = city_levels), y = mean, fill = variable)) +
  geom_col(position = "identity") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
  geom_hline(yintercept = 0, color = "black") +
  facet_wrap(~ variable_group, scales = "free_y",
             labeller = as_labeller(c(
               "SupplyDemand" = "Supply (+) vs Demand (–)",
               "UGEI_50"      = "Equity Index"
             ))) +
  scale_fill_manual(values = c(
    "BAF2_50" = "#407ca2",  # Supply
    "risk50"    = "#bc3e45",  # Demand
    "UGEI_50"   = "#855a6e"   # Equity
  )) +
  scale_x_discrete(labels = city_labels, drop = FALSE) +
  coord_cartesian(ylim = c(-max_abs_padded, max_abs_padded)) +
  labs(y = "Mean ± SD", x = "City", title = "B. Potential Supply vs Demand and Equity Index") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

```{r}
library(patchwork)

# Add titles directly when arranging
combined_plot <- a + b + plot_layout(ncol = 1)

combined_plot

ggsave("C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Paper/Figures/combined_stackplot.png", combined_plot, width = 7, height = 10, dpi = 300)
```

```{r}
# Prepare long data
summary_long2 <- summary_stats %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  # Keep only Contribution efficiency index
  filter(variable == "BAFper_50")

# Add spacer row for Manchester
spacer <- summary_long2 %>%
  distinct(variable) %>%
  mutate(city = "Spacer", mean = NA_real_, sd = NA_real_)

summary_long2 <- bind_rows(summary_long2, spacer)

# Define city order
city_levels <- c(
  "AalstADPgrdns","AntwerpenADPgrdns","BruggeADPgrdns","GenkADPgrdns","GentADPgrdns",
  "HasseltADPgrdns","KortrijkADPgrdns","LeuvenADPgrdns","MechelenADPgrdns","OostendeADPgrdns",
  "RoeselareADPgrdns","Sint","TurnhoutADPgrdns",
  "Spacer",
  "Manchester"
)

# City labels
city_labels <- c(
  AalstADPgrdns = "Aalst",
  AntwerpenADPgrdns = "Antwerpen",
  BruggeADPgrdns = "Brugge",
  GenkADPgrdns ="Genk",
  GentADPgrdns ="Gent",
  HasseltADPgrdns ="Hasselt",
  KortrijkADPgrdns ="Kortrijk",
  LeuvenADPgrdns ="Leuven",
  MechelenADPgrdns="Mechelen",
  OostendeADPgrdns ="Oostende",
  RoeselareADPgrdns ="Roeselare",
  Sint = "Sint-Niklaas",
  TurnhoutADPgrdns ="Turnhout",
  Spacer = " ",       # empty label for gap
  Manchester ="Manchester"
)

# Symmetric y-axis (optional for this single variable)
max_val <- max(summary_long2$mean + summary_long2$sd, na.rm = TRUE)
max_val_padded <- max_val * 1.05

# Plot
ggplot(summary_long2, aes(x = factor(city, levels = city_levels), y = mean)) +
  geom_col(fill = "#407ca2") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +
  scale_x_discrete(labels = city_labels, drop = FALSE) +
  coord_cartesian(ylim = c(0, max_val_padded)) +
  labs(y = "Contribution efficiency index ± SD", x = "City",
       title = "Supply: Contribution Efficiency Index per City") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}

# ---- Prep data ----
summary_long <- summary_stats %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value)

# Variable labels
summary_long <- summary_long %>%
  mutate(variable_label = recode(
    variable,
    "BAFtot_50" = "Supply",
    "risk50"    = "Demand",
    "UGEI_50"   = "Equity Index"
  )) %>%
  mutate(variable_label = factor(variable_label,
                                 levels = c("Supply", "Demand", "Equity Index")))

# ---- Ordering + labels ----
city_levels <- c(
  "AalstADPgrdns",
  "AntwerpenADPgrdns",
  "BruggeADPgrdns",
  "GenkADPgrdns",
  "GentADPgrdns",
  "HasseltADPgrdns",
  "KortrijkADPgrdns",
  "LeuvenADPgrdns",
  "MechelenADPgrdns",
  "OostendeADPgrdns",
  "RoeselareADPgrdns",
  "Sint",
  "TurnhoutADPgrdns",   # last Flemish city
  "Spacer",             # white gap
  "Manchester"          # separated at the end
)

city_labels <- c(
  AalstADPgrdns    = "Aalst",
  AntwerpenADPgrdns= "Antwerpen",
  BruggeADPgrdns   = "Brugge",
  GenkADPgrdns     = "Genk",
  GentADPgrdns     = "Gent",
  HasseltADPgrdns  = "Hasselt",
  KortrijkADPgrdns = "Kortrijk",
  LeuvenADPgrdns   = "Leuven",
  MechelenADPgrdns = "Mechelen",
  OostendeADPgrdns = "Oostende",
  RoeselareADPgrdns= "Roeselare",
  Sint             = "Sint-Niklaas",
  TurnhoutADPgrdns = "Turnhout",
  Spacer           = " ",          # single space -> keeps slot visible
  Manchester       = "Manchester"
)

# ---- Ensure Spacer rows exist for each variable ----
var_levels <- summary_long %>% distinct(variable, variable_label)
spacers <- var_levels %>% mutate(city = "Spacer", mean = NA_real_, sd = NA_real_)
summary_long <- bind_rows(summary_long, spacers)

# Make sure Spacer tiles render white
summary_long <- summary_long %>%
  mutate(mean = ifelse(city == "Spacer", NA_real_, mean))

# ---- Combined heatmap (3 rows, shared x) ----
ggplot() +
  # Supply
  geom_tile(
    data = summary_long %>% filter(variable == "BAFtot_50"),
    aes(x = city, y = variable_label, fill = mean),
    color = "white"
  ) +
  scale_fill_gradient(name = "Supply", low = "#e8e8e8", high = "#407ca2", na.value = "white") +
  new_scale_fill() +
  # Demand
  geom_tile(
    data = summary_long %>% filter(variable == "risk50"),
    aes(x = city, y = variable_label, fill = mean),
    color = "white"
  ) +
  scale_fill_gradient(name = "Demand", low = "#e8e8e8", high = "#bc3e45", na.value = "white") +
  new_scale_fill() +
  # Equity Index
  geom_tile(
    data = summary_long %>% filter(variable == "UGEI_50"),
    aes(x = city, y = variable_label, fill = mean),
    color = "white"
  ) +
  scale_fill_gradient(name = "Equity Index", low = "#bc3e45", high = "#407ca2", na.value = "white") +
  scale_x_discrete(limits = city_levels, labels = city_labels, drop = FALSE) +
  labs(x = "City", y = "", title = "Garden Metrics Heatmap") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10)
  )

# ---- Patchwork version (stacked) ----
heatmap_single <- function(var_name, title, low_col, high_col){
  ggplot(
    summary_long %>% filter(variable_label == var_name),
    aes(x = city, y = variable_label, fill = mean)
  ) +
    geom_tile(color = "white") +
    scale_fill_gradient(name = title, low = low_col, high = high_col, na.value = "white") +
    scale_x_discrete(limits = city_levels, labels = city_labels, drop = FALSE) +
    labs(x = NULL, y = NULL) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_blank(),
      legend.position = "right"
    )
}

p1 <- heatmap_single("Supply", "Supply", "#e8e8e8", "#407ca2")
p2 <- heatmap_single("Demand", "Demand", "#e8e8e8", "#bc3e45")
p3 <- heatmap_single("Equity Index", "Equity Index", "#bc3e45", "#407ca2")

p1 / p2 / p3




```

```{r}
#Idea for comparison Manchester with other flemish cities: 
# drop geometry before pivoting
summary_long <- summary_stats_2 %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = -city,
    names_to = c("stat", "variable"),
    names_pattern = "(mean|sd)_(.*)",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value)


# Prepare wide data without BAFper_50
summary_wide <- summary_long %>%
  filter(variable != "BAFper_50") %>%
  select(city, variable, mean) %>%
  pivot_wider(names_from = variable, values_from = mean)

# Manually rescale each indicator
#summary_scaled <- summary_wide %>%
#  mutate(
#    BAFtot_50  = rescale(BAFtot_50, to = c(0,1)),       # Supply
#    risk50   = rescale(risk50, to = c(0,1)),        # Demand
#    UGEI_50       = scales::rescale(UGEI_50, to = c(-1,1))    # Equity
#  )

summary_scaled <- summary_wide

# Prepare theta labels
theta_labels <- c(
  "BAFtot_50" = "Supply",
  "risk50"  = "Demand",
  "UGEI_50"      = "Equity Index"
)

# Example: Manchester + first Flemish city
plot_ly(type = 'scatterpolar', mode = 'lines+markers') %>%
  add_trace(
    r = as.numeric(summary_scaled[1, -1]),
    theta = theta_labels[names(summary_scaled)[-1]],
    fill = 'toself',
    name = summary_scaled$city[1]
  ) %>%
  add_trace(
    r = as.numeric(summary_scaled[2, -1]),
    theta = theta_labels[names(summary_scaled)[-1]],
    fill = 'toself',
    name = summary_scaled$city[2]
  ) %>%
  layout(
    polar = list(
      radialaxis = list(visible = TRUE, range = c(-1, 1))  # Set full scale for all indicators
    ),
    showlegend = TRUE
  )



```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


all_gardens_df <- st_drop_geometry(all_gardens_15)

all_gardens_df <- all_gardens_df %>%
  rename(
    risk_50 = risk50,
    risk_100 = risk100,
    risk_200 = risk200,
    risk_400 = risk400
  )

metric_cols <- c(
  "BAF1_50","BAF1_100","BAF1_200","BAF1_400",
  "risk_50","risk_100","risk_200","risk_400",
  "UGEI_50h", "UGEI_100h", "UGEI_200h", "UGEI_400h"
)


all_long <- all_gardens_df %>%
  select(city, all_of(metric_cols)) %>%
  pivot_longer(
    cols = -city,
    names_to = c("variable", "buffer"),
    names_pattern = "(.*)_(\\d+)",
    values_to = "value"
  ) %>%
  mutate(buffer = as.numeric(buffer))

summary_all <- all_long %>%
  group_by(variable, buffer) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE),
    .groups = "drop"
  )

# Create a named vector for nicer labels
variable_labels <- c(
  BAF1 = "Supply index",
  risk = "Demand index",
  UGEI = "Equity index"
)

# Assuming summary_all is ready from your previous code

# Define colors for variables
variable_colors <- c(
  "BAF1" = "#407ca2",  # Supply
  "risk"   = "#bc3e45",  # Demand
  "UGEI"   = "#855a6e"   # Equity
)

ggplot(summary_all, aes(x = buffer, y = mean_value, color = variable, fill = variable)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
              alpha = 0.2, color = NA) +
  facet_wrap(~ variable, scales = "free_y",
             labeller = labeller(variable = variable_labels)) +
  scale_color_manual(values = variable_colors) +
  scale_fill_manual(values = variable_colors) +
  scale_x_continuous(
    breaks = c(50, 100, 200, 400),
    labels = c("5,000 m²", "25,000 m²", "100,000 m²", "500,000 m²")
  ) +
  labs(
    x = "Buffer area",
    y = "Mean metric value ± SD",
    title = "A. Flanders - Average Garden Metrics"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "none",
    panel.grid.minor = element_blank()
  )


```
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


all_gardens_df <- st_drop_geometry(Manchester_gardens)

all_gardens_df <- all_gardens_df %>%
  rename(
    risk_50 = risk50,
    risk_100 = risk100,
    risk_200 = risk200,
    risk_400 = risk400
  )

metric_cols <- c(
  "BAF1_50","BAF1_100","BAF1_200","BAF1_400",
  "risk_50","risk_100","risk_200","risk_400",
  "UGEI_50h", "UGEI_100h", "UGEI_200h", "UGEI_400h"
)


all_long <- all_gardens_df %>%
  select(city, all_of(metric_cols)) %>%
  pivot_longer(
    cols = -city,
    names_to = c("variable", "buffer"),
    names_pattern = "(.*)_(\\d+)",
    values_to = "value"
  ) %>%
  mutate(buffer = as.numeric(buffer))

summary_all <- all_long %>%
  group_by(variable, buffer) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE),
    .groups = "drop"
  )

# Create a named vector for nicer labels
variable_labels <- c(
  BAF1 = "Supply index",
  risk = "Demand index",
  UGEI = "Equity index"
)

# Assuming summary_all is ready from your previous code

# Define colors for variables
variable_colors <- c(
  "BAF1" = "#407ca2",  # Supply
  "risk"   = "#bc3e45",  # Demand
  "UGEI"   = "#855a6e"   # Equity
)

ggplot(summary_all, aes(x = buffer, y = mean_value, color = variable, fill = variable)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
              alpha = 0.2, color = NA) +
  facet_wrap(~ variable, scales = "free_y",
             labeller = labeller(variable = variable_labels)) +
  scale_color_manual(values = variable_colors) +
  scale_fill_manual(values = variable_colors) +
  scale_x_continuous(
    breaks = c(50, 100, 200, 400),
    labels = c("5,000 m²", "25,000 m²", "100,000 m²", "500,000 m²")
  ) +
  labs(
    x = "Buffer area",
    y = "Mean metric value ± SD",
    title = "B. Manchester - Average Garden Metrics"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "none",
    panel.grid.minor = element_blank()
  )


```


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Define colors and linetypes
variable_colors <- c(
  "BAFtot" = "#407ca2",  # Supply
  "risk"   = "#bc3e45",  # Demand
  "UGEI"   = "#855a6e"   # Equity
)
variable_linetypes <- c(
  "BAFtot" = "solid",
  "risk"   = "dashed",
  "UGEI"   = "dotdash"
)

ggplot(summary_all, aes(x = buffer, y = mean_value, color = variable, fill = variable, linetype = variable)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
              alpha = 0.2, color = NA) +
  scale_color_manual(values = variable_colors, labels = variable_labels) +
  scale_fill_manual(values = variable_colors, labels = variable_labels) +
  scale_linetype_manual(values = variable_linetypes, labels = variable_labels) +
  scale_x_continuous(
    breaks = c(50, 100, 200, 400),
    labels = c("5,000 m²", "25,000 m²", "100,000 m²", "500,000 m²")
  ) +
  labs(
    x = "Buffer area",
    y = "Mean metric value ± SD",
    color = "Metric",
    fill = "Metric",
    linetype = "Metric",
    title = "A. Average Garden Metrics - Flanders"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    legend.position = "top",
    panel.grid.minor = element_blank()
  )
```

```{r}
# List of folders for each year
years <- c("BBK_12", "BBK_15", "BBK_18", "BBK_21")
base_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1"

# --- Function to read shapefiles, compute indicators, summarise per city ---
process_year <- function(year_folder) {
  shp_dir <- file.path(base_dir, year_folder)
  shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)
  
  # Read shapefile and add city column
  read_with_city <- function(file) {
    city <- str_extract(basename(file), "^[A-Za-z]+")
    st_read(file, quiet = TRUE) %>%
      mutate(city = city)
  }
  
  # Combine all shapefiles for the year
  all_gardens <- map(shp_files, read_with_city) %>% bind_rows()
  
  # Filter and compute UGEI for multiple distances
  all_gardens <- all_gardens %>%
    filter(ex50 != 0) %>%
    mutate(
      UGEI_50  = BAFtot_50  - risk50_2,
      UGEI_100 = BAFtot_100 - risk100_2,
      UGEI_200 = BAFtot_200 - risk200_2,
      UGEI_400 = BAFtot_400 - risk400_2
    )
  
  # Summarise indicators per city
  summary_stats <- all_gardens %>%
    group_by(city) %>%
    summarise(
      mean_UGEI_50   = mean(UGEI_50, na.rm = TRUE),
      mean_BAFtot_50 = mean(BAFtot_50, na.rm = TRUE),
      mean_BAFper_50 = mean(BAFper_50, na.rm = TRUE),
      mean_risk50_2  = mean(risk50_2, na.rm = TRUE),
      mean_HG_50     = mean(HG50, na.rm = TRUE),
      mean_W_50      = mean(W50, na.rm = TRUE),
      mean_BS_50     = mean(BS50, na.rm = TRUE),
      mean_LG_50     = mean(LG50, na.rm = TRUE),
      mean_A_50      = mean(A50, na.rm = TRUE),
      sum_HG_50     = sum(HG50, na.rm = TRUE),
      sum_W_50      = sum(W50, na.rm = TRUE),
      sum_BS_50     = sum(BS50, na.rm = TRUE),
      sum_LG_50     = sum(LG50, na.rm = TRUE),
      sum_A_50      = sum(A50, na.rm = TRUE),
      mean_HGsur_50     = mean(HGsur50, na.rm = TRUE),
      mean_Wsur_50      = mean(Wsur50, na.rm = TRUE),
      mean_BSsur_50     = mean(BSsur50, na.rm = TRUE),
      mean_LGsur_50     = mean(LGsur50, na.rm = TRUE),
      mean_Asur_50      = mean(Asur50, na.rm = TRUE),
      sum_HGsur_50     = sum(HGsur50, na.rm = TRUE),
      sum_Wsur_50      = sum(Wsur50, na.rm = TRUE),
      sum_BSsur_50     = sum(BSsur50, na.rm = TRUE),
      sum_LGsur_50     = sum(LGsur50, na.rm = TRUE),
      sum_Asur_50      = sum(Asur50, na.rm = TRUE)
    ) %>%
    mutate(
      year = as.numeric(str_extract(year_folder, "\\d+"))
    )
  
  return(summary_stats)
}

# --- Run for all years and combine ---
all_years <- map_dfr(years, process_year)


# Quick check
head(all_years)
```

```{r}
city_labels <- c(
  AalstADPgrdns = "Aalst",
  AntwerpenADPgrdns = "Antwerpen",
  BruggeADPgrdns = "Brugge",
  GenkADPgrdns ="Genk",
  GentADPgrdns ="Gent",
  HasseltADPgrdns ="Hasselt",
  KortrijkADPgrdns ="Kortrijk",
  LeuvenADPgrdns ="Leuven",
  MechelenADPgrdns="Mechelen",
  OostendeADPgrdns ="Oostende",
  RoeselareADPgrdns ="Roeselare",
  Sint = "Sint-Niklaas",
  TurnhoutADPgrdns ="Turnhout"
  # add all your cities here
)

# Plot temporal evolution (example: mean_UGEI per city)
library(ggplot2)
library(viridis)   # for nice colors

# Plot
p <- ggplot(all_years, aes(x = year, y = mean_UGEI_50, color = city, group = city)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_viridis_d(labels = city_labels, breaks = names(city_labels)) +
  scale_x_continuous(breaks = c(2012, 2015, 2018, 2021)) +
  labs(
    x = "Year",
    y = "Mean UGEI (50 m buffer)",
    color = "City",
    title = "Temporal Evolution of UGEI (50 m) per City"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11)
  )

# Show plot
p

# Optionally save
#ggsave("UGEI_temporal_evolution.png", plot = p, width = 10, height = 6, dpi = 300)

# --- average across all cities ---
all_years_avg <- all_years %>%
  group_by(year) %>%
  summarise(mean_UGEI_50 = mean(mean_UGEI_50, na.rm = TRUE))

# --- plot ---
p_avg <- ggplot(all_years_avg, aes(x = year, y = mean_UGEI_50)) +
  geom_line(color = "darkgreen", size = 1.2) +
  geom_point(color = "forestgreen", size = 3) +
  scale_x_continuous(breaks = c(2012, 2015, 2018, 2021)) +
  labs(
    x = "Year",
    y = "Average Mean UGEI (50 m buffer)",
    title = "Temporal Evolution of Average UGEI (50 m) across Cities"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11)
  )

# Show
p_avg

# Optionally save
#ggsave("UGEI_temporal_evolution_avg.png", plot = p_avg, width = 8, height = 5, dpi = 300)

library(tidyverse)
library(sf)

# --- average supply & UGEI across all cities ---
all_years_avg <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    mean_BAFtot_50 = mean(mean_BAFtot_50, na.rm = TRUE),
    sd_BAFtot_50   = sd(mean_BAFtot_50, na.rm = TRUE),
    mean_UGEI_50   = mean(mean_UGEI_50, na.rm = TRUE),
    sd_UGEI_50     = sd(mean_UGEI_50, na.rm = TRUE)
  )

# --- Plot BAFtot (Supply) ---
p_BAFtot <- ggplot(all_years_avg, aes(x = year, y = mean_BAFtot_50)) +
  geom_line(color = "darkblue", size = 1.2) +
  geom_point(color = "darkblue", size = 3) +
  geom_errorbar(aes(ymin = mean_BAFtot_50 - sd_BAFtot_50, 
                    ymax = mean_BAFtot_50 + sd_BAFtot_50),
                width = 0.25, color = "grey40") +
  labs(
    x = "Year",
    y = "BAFtot (50 m) Average",
    title = "Temporal Evolution of Garden Supply (BAFtot)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

# --- Plot UGEI ---
p_UGEI <- ggplot(all_years_avg, aes(x = year, y = mean_UGEI_50)) +
  geom_line(color = "darkgreen", size = 1.2) +
  geom_point(color = "darkgreen", size = 3) +
  geom_errorbar(aes(ymin = mean_UGEI_50 - sd_UGEI_50, 
                    ymax = mean_UGEI_50 + sd_UGEI_50),
                width = 0.25, color = "grey40") +
  labs(
    x = "Year",
    y = "UGEI (50 m) Average",
    title = "Temporal Evolution of UGEI"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

# --- Display plots individually ---
p_BAFtot
p_UGEI

library(tidyverse)

# --- Average HG, W, BS, LG, A across all cities ---
all_years_avg_extra <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    mean_HG_50 = mean(mean_HG_50, na.rm = TRUE),
    mean_W_50  = mean(mean_W_50,  na.rm = TRUE),
    mean_BS_50 = mean(mean_BS_50, na.rm = TRUE),
    mean_LG_50 = mean(mean_LG_50, na.rm = TRUE),
    mean_A_50  = mean(mean_A_50,  na.rm = TRUE)
  )

# --- Reshape to long format for plotting ---
all_years_long <- all_years_avg_extra %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(indicator = recode(indicator,
                            mean_HG_50 = "HG",
                            mean_W_50  = "W",
                            mean_BS_50 = "BS",
                            mean_LG_50 = "LG",
                            mean_A_50  = "A"))

# --- Plot all in one graph ---
p_indicators <- ggplot(all_years_long, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(2012, 2015, 2018, 2021)) +
  labs(
    x = "Year",
    y = "Average (50 m buffer)",
    color = "Indicator",
    title = "Temporal Evolution of Garden Indicators (HG, W, BS, LG, A)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold")
  )

# Show plot
p_indicators

# --- Average HG, W, BS, LG, A across all cities ---
all_years_avg_extra <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    sum_HG_50 = sum(sum_HG_50, na.rm = TRUE),
    sum_W_50  = sum(sum_W_50,  na.rm = TRUE),
    sum_BS_50 = sum(sum_BS_50, na.rm = TRUE),
    sum_LG_50 = sum(sum_LG_50, na.rm = TRUE),
    sum_A_50  = sum(sum_A_50,  na.rm = TRUE)
  )

# --- Reshape to long format for plotting ---
all_years_long <- all_years_avg_extra %>%
  pivot_longer(cols = starts_with("sum_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(indicator = recode(indicator,
                            sum_HG_50 = "HG",
                            sum_W_50  = "W",
                            sum_BS_50 = "BS",
                            sum_LG_50 = "LG",
                            sum_A_50  = "A"))

# --- Plot all in one graph ---
p_indicators <- ggplot(all_years_long, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(2012, 2015, 2018, 2021)) +
  labs(
    x = "Year",
    y = "Total",
    color = "Indicator",
    title = "Temporal Evolution of Garden Indicators (HG, W, BS, LG, A)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold")
  )

# Show plot
p_indicators

```

```{r}


# --- Example: Temporal evolution of mean UGEI per city ---
ggplot(all_years, aes(x = year, y = mean_UGEI_50, color = city, group = city)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_viridis_d(labels = city_labels, breaks = names(city_labels)) +
  scale_x_continuous(breaks = c(2012, 2015, 2018, 2021)) +
  labs(
    x = "Year",
    y = "Mean UGEI (50 m buffer)",
    color = "City",
    title = "Temporal Evolution of UGEI per City"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11)
  )

# --- Temporal evolution of Supply (BAFtot) and UGEI averages across cities ---
all_years_avg <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    mean_BAFtot_50 = mean(mean_BAFtot_50, na.rm = TRUE),
    sd_BAFtot_50   = sd(mean_BAFtot_50, na.rm = TRUE),
    mean_UGEI_50   = mean(mean_UGEI_50, na.rm = TRUE),
    sd_UGEI_50     = sd(mean_UGEI_50, na.rm = TRUE)
  )

# Supply (BAFtot)
p_BAFtot <- ggplot(all_years_avg, aes(x = year, y = mean_BAFtot_50)) +
  geom_line(color = "#407ca2", size = 1.3) +
  geom_point(color = "#407ca2", size = 3) +
  geom_ribbon(aes(ymin = mean_BAFtot_50 - sd_BAFtot_50,
                  ymax = mean_BAFtot_50 + sd_BAFtot_50),
              alpha = 0.2, fill = "#407ca2") +
  scale_x_continuous(breaks = unique(all_years_avg$year)) +
  labs(x = "Year", y = "Supply (5000 m²) Average",
       title = "Temporal Evolution of Garden Supply") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))

# UGEI
p_UGEI <- ggplot(all_years_avg, aes(x = year, y = mean_UGEI_50)) +
  geom_line(color = "#8b5fbf", size = 1.3) +
  geom_point(color = "#8b5fbf", size = 3) +
  geom_ribbon(aes(ymin = mean_UGEI_50 - sd_UGEI_50,
                  ymax = mean_UGEI_50 + sd_UGEI_50),
              alpha = 0.2, fill = "#8b5fbf") +
  scale_x_continuous(breaks = unique(all_years_avg$year)) +
  labs(x = "Year", y = "UGEI (5000 m²) Average",
       title = "Temporal Evolution of Garden Equity Index") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))

p_BAFtot
p_UGEI


# --- Combined plot for indicators HG, W, BS, LG, A ---

# Reshape and recode
all_years_long <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    mean_HG_50 = mean(mean_HG_50, na.rm = TRUE),
    mean_LG_50 = mean(mean_LG_50, na.rm = TRUE),
    mean_W_50  = mean(mean_W_50, na.rm = TRUE),
    mean_A_50  = mean(mean_A_50, na.rm = TRUE),
    mean_BS_50 = mean(mean_BS_50, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(
    indicator = recode(indicator,
                       mean_HG_50 = "High green",
                       mean_LG_50 = "Low green",
                       mean_W_50  = "Water",
                       mean_A_50  = "Sealed surface",
                       mean_BS_50 = "Bare soil")
  )

# Define custom colors
indicator_colors <- c(
  "High green" = "#607e32",
  "Low green"  = "#a7bc7a",
  "Water"      = "#3f70a8",
  "Sealed surface" = "#8e8e8e",
  "Bare soil"  = "#9b875b"
)

# Plot
p_indicators <- ggplot(all_years_long, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = indicator_colors) +
  scale_x_continuous(breaks = unique(all_years_long$year)) +
  labs(
    x = "Year",
    y = "Average across cities",
    color = "Indicator",
    title = "Temporal Evolution of Garden Indicators"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11),
    legend.position = "top"
  )

library(dplyr)
library(tidyr)
library(ggplot2)

# --- Summed indicators across all cities ---
all_years_sum <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    sum_HG_50 = sum(sum_HG_50, na.rm = TRUE),
    sum_LG_50 = sum(sum_LG_50, na.rm = TRUE),
    sum_W_50  = sum(sum_W_50, na.rm = TRUE),
    sum_A_50  = sum(sum_A_50, na.rm = TRUE),
    sum_BS_50 = sum(sum_BS_50, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("sum_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(indicator = recode(indicator,
                            sum_HG_50 = "High green",
                            sum_LG_50 = "Low green",
                            sum_W_50  = "Water",
                            sum_A_50  = "Sealed surface",
                            sum_BS_50 = "Bare soil"))

# Define colors (same as before)
indicator_colors <- c(
  "High green" = "#607e32",
  "Low green"  = "#a7bc7a",
  "Water"      = "#3f70a8",
  "Sealed surface" = "#8e8e8e",
  "Bare soil"  = "#9b875b"
)

# --- Plot sum ---
p_indicators_sum <- ggplot(all_years_sum, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = indicator_colors) +
  scale_x_continuous(breaks = unique(all_years_sum$year)) +
  labs(
    x = "Year",
    y = "Sum across cities",
    color = "Indicator",
    title = "Temporal Evolution of Total Garden Indicators"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11),
    legend.position = "top"
  )



# Reshape and recode
all_years_long <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    mean_HGsur_50 = mean(mean_HGsur_50, na.rm = TRUE),
    mean_LGsur_50 = mean(mean_LGsur_50, na.rm = TRUE),
    mean_Wsur_50  = mean(mean_Wsur_50, na.rm = TRUE),
    mean_Asur_50  = mean(mean_Asur_50, na.rm = TRUE),
    mean_BSsur_50 = mean(mean_BSsur_50, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(
    indicator = recode(indicator,
                       mean_HGsur_50 = "High green",
                       mean_LGsur_50 = "Low green",
                       mean_Wsur_50  = "Water",
                       mean_Asur_50  = "Sealed surface",
                       mean_BSsur_50 = "Bare soil")
  )

# Define custom colors
indicator_colors <- c(
  "High green" = "#607e32",
  "Low green"  = "#a7bc7a",
  "Water"      = "#3f70a8",
  "Sealed surface" = "#8e8e8e",
  "Bare soil"  = "#9b875b"
)

# Plot
p_indicators_sur <- ggplot(all_years_long, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = indicator_colors) +
  scale_x_continuous(breaks = unique(all_years_long$year)) +
  labs(
    x = "Year",
    y = "Average across cities",
    color = "Indicator",
    title = "Temporal Evolution of Garden + Buffer Indicators"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11),
    legend.position = "top"
  )


# --- Summed indicators across all cities ---
all_years_sum <- all_years %>%
  st_drop_geometry() %>%
  group_by(year) %>%
  summarise(
    sum_HGsur_50 = sum(sum_HGsur_50, na.rm = TRUE),
    sum_LGsur_50 = sum(sum_LGsur_50, na.rm = TRUE),
    sum_Wsur_50  = sum(sum_Wsur_50, na.rm = TRUE),
    sum_Asur_50  = sum(sum_Asur_50, na.rm = TRUE),
    sum_BSsur_50 = sum(sum_BSsur_50, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("sum_"),
               names_to = "indicator",
               values_to = "value") %>%
  mutate(indicator = recode(indicator,
                            sum_HGsur_50 = "High green",
                            sum_LGsur_50 = "Low green",
                            sum_Wsur_50  = "Water",
                            sum_Asur_50  = "Sealed surface",
                            sum_BSsur_50 = "Bare soil"))

# Define colors (same as before)
indicator_colors <- c(
  "High green" = "#607e32",
  "Low green"  = "#a7bc7a",
  "Water"      = "#3f70a8",
  "Sealed surface" = "#8e8e8e",
  "Bare soil"  = "#9b875b"
)

# --- Plot sum ---
p_indicators_sum_sur <- ggplot(all_years_sum, aes(x = year, y = value, color = indicator, group = indicator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = indicator_colors) +
  scale_x_continuous(breaks = unique(all_years_sum$year)) +
  labs(
    x = "Year",
    y = "Sum surroundings across cities",
    color = "Indicator",
    title = "Temporal Evolution of Total Garden + Buffer Indicators"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 11),
    legend.position = "top"
  )




# Show plots
p_BAFtot
p_UGEI
p_indicators
p_indicators_sum
p_indicators_sur
p_indicators_sum_sur

```

```{r}
# path to folder
shp_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/BBK_15"

# list shapefiles
shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)

# function to read shapefile and add city column
read_with_city <- function(file) {
  city <- str_extract(basename(file), "^[A-Za-z]+")  # take leading word (e.g. Aalst)
  st_read(file, quiet = TRUE) %>%
    mutate(city = city)
}

# read all and bind
all_gardens_15 <- map(shp_files, read_with_city) %>%
  bind_rows()
```

#probleem met tuinen die op de rand van twee rasters vallen? of waarvan 400 m erbuiten niet meer binnen raster valt?
```{r}
all_gardens_15 <- all_gardens_15 %>% filter(ex50 != 0)

sum(duplicated(all_gardens_15$CAPAKEY))
sum(all_gardens_15$sur50>6000)
sum(all_gardens_15$sur50<4000)

test <- all_gardens_15 %>% filter(sur50 < 4000 & city == 'LeuvenADPgrdns')
```

```{r}
# path to folder
shp_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/BBK_18"

# list shapefiles
shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)

# function to read shapefile and add city column
read_with_city <- function(file) {
  city <- str_extract(basename(file), "^[A-Za-z]+")  # take leading word (e.g. Aalst)
  st_read(file, quiet = TRUE) %>%
    mutate(city = city)
}

# read all and bind
all_gardens_18 <- map(shp_files, read_with_city) %>%
  bind_rows()
```

#probleem met tuinen die op de rand van twee rasters vallen? of waarvan 400 m erbuiten niet meer binnen raster valt?
```{r}
all_gardens_18 <- all_gardens_18 %>% filter(ex50 != 0)

sum(duplicated(all_gardens_18$CAPAKEY))
sum(all_gardens_18$sur50>6000)
sum(all_gardens_18$sur50<4000)
```

```{r}
# path to folder
shp_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/BBK_12"

# list shapefiles
shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)

# function to read shapefile and add city column
read_with_city <- function(file) {
  city <- str_extract(basename(file), "^[A-Za-z]+")  # take leading word (e.g. Aalst)
  st_read(file, quiet = TRUE) %>%
    mutate(city = city)
}

# read all and bind
all_gardens_12 <- map(shp_files, read_with_city) %>%
  bind_rows()
```

#probleem met tuinen die op de rand van twee rasters vallen? of waarvan 400 m erbuiten niet meer binnen raster valt?
```{r}
all_gardens_12 <- all_gardens_12 %>% filter(ex50 != 0)

sum(duplicated(all_gardens_12$CAPAKEY))
sum(all_gardens_12$sur50>6000)
sum(all_gardens_12$sur50<4000)
```


```{r}
# path to folder
shp_dir <- "C:/Users/u0137615/OneDrive - KU Leuven/Phd GARLOCK/3. Onderzoek/WP3/Output/Garden_metric1/BBK_21"

# list shapefiles
shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)

# function to read shapefile and add city column
read_with_city <- function(file) {
  city <- str_extract(basename(file), "^[A-Za-z]+")  # take leading word (e.g. Aalst)
  st_read(file, quiet = TRUE) %>%
    mutate(city = city)
}

# read all and bind
all_gardens_21 <- map(shp_files, read_with_city) %>%
  bind_rows()
```

#probleem met tuinen die op de rand van twee rasters vallen? of waarvan 400 m erbuiten niet meer binnen raster valt?
```{r}
all_gardens_21 <- all_gardens_21 %>% filter(ex50 != 0)

sum(duplicated(all_gardens_21$CAPAKEY))
sum(all_gardens_21$sur50>6000)
sum(all_gardens_21$sur50<4000)
```
